PHYS213 Paper

### 2022-08-04 latex/PHYS213/chapters/chapter1.tex
add LaTeX files

\documentclass[a4paper]{report}

\usepackage{needed}

\title{PHYS213}
\author{Alfaifi, Ammar}
\date{\today}

\begin{document}
    \maketitle
    \chapter{The Quantum Theory of Light}
        \section{Heartz's Experiments; Light as EM Wave}

@@

            \begin{wrapfigure}{r}{0.2\linewidth}
                \includegraphics[width=0.9\linewidth]{figures/x-ray acattering.png}
                \caption{Bragg scattering of x-rays from successive planes of atoms.}
                \label{fig:Bragg scattering}
            \end{wrapfigure}

@@

            \paragraph{$\bullet$} The adjacent atoms in a single plane, $A$, will scatter constructively if the angle of 
            incidence, $\theta_i$, equals angel of reflaction, $\theta_r$.\\
            Atoms in \textit{successive planes} (A \& B), will scatter constructively if the the path length difference for rays (1) \& (2)
            is a ehole number of wavelength, thus $n\theta$, so 
            \begin{align}
                \label{The condition of scattering}
                \overline{AB} + \overline{BC} = n\theta, \qquad n=1, 2, 3, \dots
            \end{align}
            Thus Bragg equation is 
            \coloredeq{eq:Bragg equation}{\Delta{L} = 2d\sin{\theta} = n\lambda}

\end{document}

>>>>

\chapter{The Quantum Theory of Light}
\section{Heartz's Experiments; Light as EM Wave}

@@

    \begin{wrapfigure}{r}{0.2\linewidth}
        \includegraphics[width=0.9\linewidth]{figures/x-ray acattering.png}
        \caption{Bragg scattering of x-rays from successive planes of atoms.}
        \label{fig:Bragg scattering}
        \includegraphics[width=0.9\linewidth]{figures/x-ray spectrum.png}
        \caption{The x-ray spectrum of a metal target.}
        \label{fig:x-ray spectrum}
    \end{wrapfigure}

@@

    \paragraph{$\bullet$} The adjacent atoms in a single plane, $A$, will scatter constructively if the angle of 
    incidence, $\theta_i$, equals angel of reflaction, $\theta_r$.\\
    Atoms in \textit{successive planes} (A \& B), will scatter constructively if the the path length difference for rays (1) \& (2)
    is a ehole number of wavelength, thus $n\theta$, so 
    \begin{align}
        \label{The condition of scattering}
        \overline{AB} + \overline{BC} = n\theta, \qquad n=1, 2, 3, \dots
    \end{align}
    Thus Bragg equation is 
    \coloredeq{eq:Bragg equation}{\Delta{L} = 2d\sin{\theta} = n\lambda}
    {\tiny \begin{itemize}
        \item $n$ is the order of the intensity maximum, 
        \item $\lambda$ is the x-ray wavelength, 
        \item $d$ is the spacing between planes, 
        \item $\theta$ is the angle of the intensity maximum measured from plane $A$.
    \end{itemize}}
    If $\lambda$ is known $d$ of crystal can be found. If $d$ is also known, the x-ray intensity vs. wavelength may be determined
    as in Fig.\eqref{fig:x-ray spectrum}. This continuous x-ray spectrum results from glancing or indirect scattering 
    of electrons from metal atoms. In such collisions only part of the electron’s energy is converted to electromagnetic radiation.
    This radiation is called \textbf{bremsstrahlung} (German for braking radiation), which refers to the 
    radiation given off by any charged particle when it is \textit{decelerated}.

    \paragraph{$\bullet$} The $\lambda_{min}$ is independent of the target composition and only depends on the tube voltage, $V$.
    This might be explained by assuming it to the case of a \textit{head-on electron–atom collision} in which all of the 
    incident electron’s kinetic energy is converted to electromagnetic energy in the form of a single x-ray photon. For this case we have
    \begin{align}
        \label{eq:electron energy into x-ray}
        eV = hf = \frac{hc}{\lambda_{min}} \qquad or \quad \lambda_{min} = \frac{hc}{eV}
    \end{align}

    \begin{wrapfigure}{r}{0.3\linewidth}
        \includegraphics[width=0.9\linewidth]{figures/compton effect.png}
        \caption{Scattered x-ray intensity versus wavelength of Compton scattering.}
        \label{fig:Scattered x-ray}
        \includegraphics[width=0.9\linewidth]{figures/particle-like collision.png}
        \caption{The quantum view}
        \label{fig:scattered quantum x-ray}
    \end{wrapfigure}

    \paragraph{The Compton Effect} % (fold)
    \label{par:The Compton Effect}
    In 1922, Arthur Holly Compton confirmed that x-ray photons behave like particles with momentum $hf/c$.

    \paragraph{$\bullet$} classical theory predicted that incident radiation of frequency $f_0$ should accelerate an electron 
    in the direction of propagation of the incident radiation, and it should cause \textit{forced oscillations} of the electron 
    and reradiation at frequency $f'$, where $f' \le f$.\footnote{This decrease in frequency of the reradiated wave is caused 
    by a double Doppler shift, first because the electron is receding from the incident radiation, and second because the electron
    is a moving radiator as viewed from the fixed lab frame} The frequency or wavelength of the scattered radiation should depend 
    on the length of time the electron was exposed to the incident radiation as well as on the intensity of the incident radiation.

    \paragraph{$\bullet$} Compton's result was that the wavelength of the x-ray is independent of the intensity of radiation and
    exposure duration, and depends only on the scattering angle!.

    \paragraph{\color{c3}$\bigstar$} Quantum model easily explained the lower scattered frequency $f'$, where the incident photon gives 
    some its energy $hf$ and momentum to the recoiling electron.

    \paragraph{$\bullet$}  From Fig.\eqref{fig:Scattered x-ray} the shifted peak at $\lambda'$ is caused by the scattering of 
    x-rays from nearly free electrons. Assuming the photon exhibits particle-like behavior and collides elastically with a free
    electron initially at rest, Fig.\eqref{fig:scattered quantum x-ray}. Because electron recoils at high speed, we treat the
    collision relativistically. Thus from the energy conservation, 
    \begin{align}
        \label{eq:Compton derivation:1}
        E + m_ec^2 = E' + E_e
    \end{align}
    where $E$ is incident photon energy, $m_ec^2$ is the rest mass-energy of the electron, $E'$ is the energy of scattered
    photon, $E_e$ is the total relativistic energy of the electron after the collision. Likewise, by the momentum conservation, 
    \begin{align}
        \label{eq:Compton derivation:2}
        p  = p' \cos{\theta} + p_e \cos{\phi} &&
        p' \sin{\theta} = p_e \sin{\phi}
    \end{align}
    where $p$ is the momentum of the incident photon, $p'$ is the momentum of the scattered photon, and $p_e$ 
    is the recoiling electron momentum. We observed the angle of the scattered photons $\theta$, so we shall solve
    Eq.\eqref{eq:Compton derivation:1} and Eq.\eqref{eq:Compton derivation:2} simultaneously to eliminate $\phi$,
    \begin{align}
        \label{eq:Compton derivation:3}
        p_e^2 = p'^2 + p^2 - 2p p' \cos{\theta}
    \end{align}
    Paradoxically, we need to use the wave nature of light to explain the particle-like behavior 
    of photons. Recall that the energy of a photon and the frequency of the associated light wave are related 
    by $E=hf$. If we assume that a photon obeys the relativistic express $E^2 = p^2c^2 + m^2c c^4$, and the photon mass is zero, 
    \begin{align}
        \label{eq:Compton derivation:4}
        p_{photon} = \frac{E}{c} = \frac{hf}{c} = \frac{h}{\lambda}
    \end{align}
    A paradoxical situation; a particle property, the photon momentum, is given in terms of a wave property, $\lambda$, of 
    an associated light wave. If the relations $E = hf$ and $p = hf/c$ are substituted into 
    Eq.\eqref{eq:Compton derivation:1} and Eq.\eqref{eq:Compton derivation:3}, these become respectively
    \begin{align}
        \label{eq:Compton derivation:5}
        E_e = hf - hf' + m_e c^2, \qquad 
        p_e^2 = \left(\frac{hf'}{c}\right)^2 + \left(\frac{hf}{c}\right)^2 - \frac{2 h^2 f f' }{c2} \cos{\theta}
    \end{align}
    Because the Compton measurements do not concern the total energy and momentum of the electron, 
    we eliminate $E_e$ and $p_e$ by substituting Eq.\eqref{eq:Compton derivation:4} into the following expression for 
    the electron’s relativistic energy,
    \begin{align}
        \label{eq:C}
        E_e^2 = p_e^2c^2 + m_e^2 c^4
    \end{align}
    And we get the Compton's equation,
    \coloredeq{eq:Compton equation}{\lambda - \lambda' = \frac{h}{m_e c}(1-\cos{\theta})}
    {\tiny \begin{itemize}
        \item $\lambda_C = \frac{h}{m_e c}$ is Compton wavelength = \SI{0.0243}{\AA}
    \end{itemize}}
    % paragraph The Compton Effect (end)

    \paragraph{Particle-Wave Complementarity} % (fold)
    \label{par:Particle-Wave Complementarity}
    Compton effect offers evidence that when light \textit{interacts} with matter it behaves as if it were composed 
    of particles with energy $hf$ and momentum $h/\lambda$.
    % paragraph Particle-Wave Complementarity (end)
